E Coli Project Paper Incredibly Rough Draft

### 2013-09-03 paper/paper.tex
Worked on Introduction and added a lot of references to paper Tested everything on node26.

%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  This paper is about science.
\end{abstract}
\section{Introduction}
\fixme{I need a lot of references here!!!  (And the writing is just terrible)}
It is vital that the process of bacterial cell division result in a single
nucleiod in each daughter cell.  The cell must avoid minicelling, or
splitting into daughter cells with lopsided volumes.  One of the mechanisms
the cell employs in order to avoid this is to allow a long FtsZ polymer
chain to develop on the cell wall in the center of the cell.  This chain
acts as a guide for the cell splitting process.  Huang et all have shown in
simulation that a system of Min protein interaction within the cell will
lead to a natural oscillation of the MinC protein, which exhibits an
aversion to the FtsZ polymer, from one cell end to the other that will
leave the center free of any build up of MinC.  This will allow for the
PtsZ chain to develop in the center and not at the ends, where the two
nucleiods are housed.  The interaction takes place between MinE,MinD, and
MinC, which naturally associates with MinD.

\cite{howard2005cellular}\cite{kruse2002dynamic}

@@

\section{Introduction}
\fixme{I need a lot of references here!!!}
It is vital that the process of bacterial cell division result in a single
nucleiod in each daughter cell.  The cell must avoid minicelling, or
splitting into daughter cells with lopsided volumes.  One of the mechanisms
the cell employs in order to avoid this is to allow a long FtsZ polymer
chain to develop on the cell wall in the center of the cell.  This chain
acts as a guide for the cell splitting process.  Huang et all have shown in
simulation that a system of Min protein interaction within the cell will
lead to a natural oscillation of the MinC protein, which exhibits an
aversion to the FtsZ polymer, from one cell end to the other that will
leave the center free of any build up of MinC.  This will allow for the
PtsZ chain to develop in the center and not at the ends, where the two
nucleiods are housed.  The interaction takes place between MinE,MinD, and
MinC, which naturally associates with MinD.

>>>>

%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  This paper is about science.
\end{abstract}
\section{Introduction}
It is vital that during the process of bacterial cell division a
cell avoid minicelling, or splitting into daughter cells with lopsided
volumes.  During this process a long FtsZ polymer chain develops on
the cell wall in the center region of the cell which acts as a guide
for the division. Previous experimental studies have shown
that the MinC protein, known to inhibit the FtZ polymer, exhibits a
pole to pole oscillatory behavoir in conjunction with the MinD and
MinE proteins, while the oscillating MinE tends to be more localized
in the center of the cell \cite{hu1999topological} \cite{fu2001mine}
The MinC protein will then have a higher time averaged
concentration in the cell poles as opposed to the center region of the
cell, prohibiting the FtZ from building in the wrong region.

A significant amount of work has been done to develop protein reaction
and diffusion models that exhibit accurate macroscopic dynamics of the
MinD protein system. Early models involved free proteins that affect
eachothers' rates of diffusion and membrane attachement but do not
combine into compound states\cite{meinhardt2001pattern} .  In 2003
Huang \cite{huang2003dynamic} improved upon this work with a simple
model based on MinD-MinE combination, ATPase hydrolysis, and MinD
membrane attechent that when simulated exhibits accurate MinD
oscillations in cylindrical cells\cite{huang2003dynamic}. In this
model cytoplasmic MinD is more likely to attach to the membrane when
MinD is already clustered there (following observed non-linear
attachement of minD on the cell membrane), and is stationary once
attached.  A number of studies have used an approach similar in that
they do not rely on the ability of MinD to move along the walls and
cluster ~\cite{kruse2007experimentalist} \cite{meinhardt2001pattern}
\cite{drew2005polymerization} \cite{fange2006noise}, while studies
have been made as well of models which rely on MinD mobility and
attraction on the cell
membrane\cite{kruse2002dynamic}\cite{howard2005cellular}.  Variations
of the Huang 2003 model that attempt to stochasticly account for
variations of molecular interaction confirm the major results obtained
by Huang's model and more successfully predict experimentally observed
oscillations in round cell phenotypes.\cite{drew2005polymerization}
\cite{fange2006noise} \cite{huang2004min} Biochemical models of
broader scope have also been used to study the MinD system and show
consistent results.\cite{arjunan2010new}

Previous studies have been made of the Min system's associtation with
the cell membrane
\cite{hsieh2010direct}\cite{mileykovskaya2003effects}.  Studies have
shown as well that MinD binds preferentially to regions enriched with
cadiolipin, an anionic phospholipid that collects on regions of high
negative curvature. This mechanism has been incorporated into other
models.\cite{drew2005polymerization}
\cite{cytrynbaum2007multistranded} \cite{renner2012mind}
\cite{renner2012mind} However, this mechanism of combined clustering,
phospholipids and MinD has not been observed in real
cells. \cite{halatek2012highly}

It is vital for the mechanism by which the cell accurately divides to
be robust and indeed it has been shown to
be\cite{touhami2006temperature}.  Studies have shown that the
oscillations can effectively find poles in abberent cell
shapes\cite{varma2008min}\cite{juarez2010changes}.  However, Mannik et
al. have shown that there are limits to this robustness. Forcing cells
into flattened, irregular cell shapes adversly effects the Min
system's ability to maintain their regular oscillatory
behavior\cite{mannik2012robustness}\cite{mannik2009bacterial}.  We
follow the experimental work of Mannik et and use Huang's mean field
differential equation reaction model to explore the model's ability to
find regular oscillatory behavoir in a series of abberent shapes.

@@

\section{Introduction}
\fixme{I need a lot of references here!!!}
 need a lot of references here!!!  (And the writing is just
terrible) It is vital that during the process of bacterial cell
division that the cell avoid minicelling, or splitting into daughter
cells with lopsided volumes.  During this process a long FtsZ polymer
chain develops on the cell wall in the center region of the cell
acting as a guide for the cell splitting process.  in simulation that a system of Min protein interaction within
the cell will lead to a natural oscillation of the MinC protein, which
exhibits an aversion to the FtsZ polymer, from one cell end to the
other that will leave the center free of any build up of MinC.  This
will allow for the PtsZ chain to develop in the center and not at the
ends, where the two nucleiods are housed.  The interaction takes place
between MinE,MinD, and MinC, which naturally associates with MinD.

It has also been observed that the spatial characteristics of
oscillation is not effected by broad ranges of tempurature changes
(although the frequency does vary) allowing for very robust cell
division that is not dettered by environmental tempurature
fluctuations. \cite{touhami2006temperature}

We wish to subject Huang's seminal model to more harsh geometrical
tests of robustness and compare results to the new experimental
results of Mannik.\cite{mannik2012robustness}

Models also introduce the affect of cardiolipin, an anionic
phospholipid that collects on regions of high negative curvature (the
membrane of the cell poles) and attracs MinD, since it had been
observed that minD binds preferetntially in regions enriched with
these phospholipids \cite{drew2005polymerization}\cite{cytrynbaum2007multistranded}. However, this
mechanism of combined clustering, phospholipids and MinD has not been
observed in real cells.  \cite{halatek2012highly}

A mathematical model that describes the building of the FtZ polymer
along the midsection cell membrane.  Solved numerically using a finite
volume method. \cite{zhang2013mathematical}

Variations of the Huang 2003 model that attempt to more accurately
account for the details of molecular interaction, focusing on the
mechanism of MinD binding to other MinD on the cell membrane.  These
models confirm the major results obtained by Huang's model, while
showing some interesting additional features.
.\cite{drew2005polymerization}\cite{fange2006noise}\fixme{this last
  reference sounds like maybe the stochastic changes they made to the
  model cause the simulations to be more accurate in certain, odd
  cases.  Should we have used this model?}

More general biochemical models have also been used to study the MinD
system and show consistent results.\cite{arjunan2010new}

Stochastic models that area based on Huangs mean-field model but take
chemical fluctuations into account in a discrete fashion show the same
results as Huang's simulations in ordinary cell shapes, and better
predict experimentally observed oscillations in round cell
shapes.\cite{fange2006noise}

Purely theoretical models have been constructed that rely on min
protein's mobility on the membranes and tendency to cluster
together.\cite{kruse2002dynamic}\cite{howard2005cellular} \fixme{read
  and write about these}

It is vital that the process of bacterial cell division result in a single
nucleiod in each daughter cell.  The cell must avoid minicelling, or
splitting into daughter cells with lopsided volumes.  One of the mechanisms
the cell employs in order to avoid this is to allow a long FtsZ polymer
chain to develop on the cell wall in the center of the cell.  This chain
acts as a guide for the cell splitting process.  Huang et all have shown in
simulation that a system of Min protein interaction within the cell will
lead to a natural oscillation of the MinC protein, which exhibits an
aversion to the FtsZ polymer, from one cell end to the other that will
leave the center free of any build up of MinC.  This will allow for the
PtsZ chain to develop in the center and not at the ends, where the two
nucleiods are housed.  The interaction takes place between MinE,MinD, and
MinC, which naturally associates with MinD.
